---
title: "Results section with merged data and year as covariate"
format: html
editor: visual
---

# Low and high quality make it High C:N and Low C:N

### importing libraries

```{r}
library(tidyverse)
library(lubridate)
library(performance)
library(broom)
library(emmeans)
library(multcomp)
library(car)
library(MASS)
library(DHARMa)
library(grDevices)
library(vegan)
library(cowplot)

# set theme
theme_set(theme_classic(base_size = 14) + theme(legend.position="none"))
```

### Importing data

```{r}
# original performance data sheet
Performance_Sheet <- read_csv("/Users/mayra/Library/CloudStorage/OneDrive-UniversityofMassachusettsBoston/Manuscripts/Performance.csv") 
# rename Diet Values

Performance_Sheet[Performance_Sheet == "High-Quality"] <- "N-Rich"
Performance_Sheet[Performance_Sheet == "Low-Quality"] <- "N-Poor"

# rename Temperature Values
Performance_Sheet[Performance_Sheet == "18C"] <- "18°C"
Performance_Sheet[Performance_Sheet == "27C"] <- "27°C"


# original color data sheet
Color_Sheet <- read_csv("/Users/mayra/Library/CloudStorage/OneDrive-UniversityofMassachusettsBoston/Manuscripts/Color.csv") 

# rename Diet Values

Color_Sheet[Color_Sheet == "High-Quality"] <- "N-Rich"
Color_Sheet[Color_Sheet == "Low-Quality"] <- "N-Poor"

# rename Temperature Values
Color_Sheet[Color_Sheet == "18C"] <- "18°C"
Color_Sheet[Color_Sheet == "27C"] <- "27°C"

#  merged performance data
merged_performance <- Performance_Sheet |> dplyr::select(ID, Diet, Temperature,`Hatch Date`,`Pupa Date`, `Death Date`, `Survived?`, `Development Time`,`Pupal Mass (g)`,`Pupal Mass no mispupation`, `Adult em. Date`, Year)
# removing spaces
merged_performance_n <- merged_performance                  
colnames(merged_performance_n) <- gsub(" ", "_", colnames(merged_performance_n))
str(merged_performance_n)
# renaming columns
merged_performance_manduca_data<- merged_performance_n  |> 
mutate(ID = as.character(ID),Hatch_Date = mdy(Hatch_Date),Pupa_Date = mdy(Pupa_Date),Death_Date = mdy(Death_Date),Adult_em._Date = mdy(Adult_em._Date),TD = paste("Fed",Diet,"at",Temperature), Growth_Rate = `Pupal_Mass_(g)`/Development_Time,Year = as.character(Year)) |> rename(Pupal_Mass = `Pupal_Mass_(g)`, Survived = `Survived?`)

# merged color data 

Color_Sheet <- Color_Sheet                  
colnames(Color_Sheet) <- gsub(" ", "_", colnames(Color_Sheet))

merged_color_manduca_data <- Color_Sheet |> dplyr::select(ID:Temperature,Year ,'Size_(px/cm)':Luminance_Variance,) |>  mutate(ID = as.character(ID), Year = as.character(Year)) |> rename("Scale" = 'Size_(px/cm)' ) |> drop_na()
# removing spaces
merged_color_manduca_data <- merged_color                 
colnames(merged_color_manduca_data) <- gsub(" ", "_", colnames(merged_color_manduca_data)) 

```

### Making Linear models

```{r}
# Fitness
### merged pupal mass
merged_pupal_lm <- lm(Pupal_Mass ~ Diet * Temperature 
                      #new added year
                      +Year
                         
                      , 
               data = merged_performance_manduca_data)

str(merged_performance_manduca_data)

### merged suvival
merged_surv_glm <- glm(Survived ~ Diet * Temperature
                      #new added year
                      +Year
                       # *Sex
                       ,
                 family = binomial(link = "logit"),
                 data = merged_performance_manduca_data)
### merged development time
merged_dev_lm <- lm(Development_Time ~ Diet * Temperature
                    # new year
                    +Year
                    ,
               data = merged_performance_manduca_data)
# color
## merged_color
merged_color_lightness_lm <- lm(Luminance_Median ~ Diet * Temperature 
                         + Scale
                         # new year
                         +Year
                        ,
               data = merged_color_manduca_data)

test<-lm(Scale~Diet * Temperature + Year, data= merged_color_manduca_data)
anova(test)
summary(test)

## merged color with mean
merged_color_lightness_mean_lm <- lm(Luminance_Mean ~ Diet * Temperature
                                     + Scale
                                     # year
                                     + Year
                                     ,
                                     data = merged_color_manduca_data)

# Growth rate
## merged_growth rate
merged_growth_rate_lm <- lm(Growth_Rate ~ Diet * Temperature
                            # year
                            +Year
                         
                        ,
               data = merged_performance_manduca_data)
```

### Results

\

#### 1 \| Performance Metrics 

Anova

Pupal mass

```{r}
# merged
Anova(merged_pupal_lm)


```

survival

```{r}
# merged
Anova(merged_surv_glm)
```

Dev time

```{r}
# merged
Anova(merged_dev_lm)
```

Growth Rate

```{r}
Anova(merged_growth_rate_lm)
```

Sidak

Pupal mass

```{r}
## merged
merged_pupal_em <- emmeans(merged_pupal_lm, specs = ~ Diet + Temperature)
merged_pupal_tk_Temp_by_Diet <- cld(merged_pupal_em, adjust = "sidak") |>
  mutate(.group = letters[as.numeric(.group)]) |>
  ggplot(mapping = aes(x =  Diet
                         , y = emmean,
                       ymin = lower.CL,
                       ymax = upper.CL,
                       colour = factor(.group))) +
  geom_pointrange() +
  geom_text(mapping = aes(label = .group),
            y = rep(1,4)) +
scale_color_manual(values  = c("black","cornflowerblue","honeydew3","snow4")) +
  facet_grid(~Temperature)
merged_pupal_tk_Temp_by_Diet
```

Survival

```{r}
## merged
merged_surv_em <- emmeans(merged_surv_glm, specs = ~ Diet + Temperature) 
merged_surv_tk_Diet_by_Temp <- cld(merged_surv_em, adjust = "sidak") |>
  mutate(.group = letters[as.numeric(.group)]) |>
  ggplot(mapping = aes(x =  Temperature
                         , y = emmean,
                      ymin = asymp.LCL,
                       ymax = asymp.UCL,
                       colour = factor(.group))) +
  geom_pointrange() +
  geom_text(mapping = aes(label = .group),
            y = rep(1,4)) +
scale_color_manual(values  = c("black","cornflowerblue","honeydew3","snow4")) +
  facet_grid(~Diet)

merged_surv_tk_Diet_by_Temp
```

Development time

```{r}
## merged
merged_dev_em <- emmeans(merged_dev_lm, specs = ~ Diet + Temperature)

merged_dev_tk_Temp_by_Diet <- cld(merged_dev_em, adjust = "sidak") |>
  mutate(.group = letters[as.numeric(.group)]) |>
  ggplot(mapping = aes(x =  Diet
                         , y = emmean,
                       ymin = lower.CL,
                       ymax = upper.CL,
                       colour = factor(.group))) +
  geom_pointrange() +
  geom_text(mapping = aes(label = .group),
            y = rep(1,4)) +
scale_color_manual(values  = c("black","cornflowerblue","honeydew3","snow4")) +
  facet_grid(~Temperature)

merged_dev_tk_Temp_by_Diet
```

Growth Rate

```{r}
merged_growth_rate_em <- emmeans(merged_growth_rate_lm, specs = ~ Diet + Temperature)

merged_growth_rate_tk_Temp_by_Diet <- cld(merged_growth_rate_em, adjust = "sidak") |>
  mutate(.group = letters[as.numeric(.group)]) |>
  ggplot(mapping = aes(x =  Diet
                         , y = emmean,
                       ymin = lower.CL,
                       ymax = upper.CL,
                       colour = factor(.group))) +
  geom_pointrange() +
  geom_text(mapping = aes(label = .group),
            y = rep(1,4)) +
scale_color_manual(values  = c("black","cornflowerblue","honeydew3","snow4")) +
  facet_grid(~Temperature)

merged_growth_rate_tk_Temp_by_Diet
```

Original plot

```{r}
# merged 
#Pupal_mass
merged_pm_plot <- ggplot(data = merged_performance_manduca_data,
                     mapping =
                       aes(x = Diet, y = Pupal_Mass_no_mispupation, fill = Temperature)) + 
  geom_boxplot() +
  facet_wrap(~Temperature)+
 scale_color_grey()  +
  scale_fill_manual(values  = c("cornsilk4","lightgray"))+
  geom_jitter(alpha = 0.5) +
  scale_y_continuous(limits =c(0,10)) +
  labs(
    #title ="Diet and Temperatures effect on Pupal Mass",
     y = "Pupal Mass(g)" , x = "Treatment" ) #+
  # trying basic sd error bars
 # geom_errorbar(data = pm_trial_1_error, aes(ymin = Mean - SD ,ymax = Mean + SD))

merged_pm_plot


# Survival
## merged
merged_surv_plot <- ggplot(data = merged_performance_manduca_data,
                     mapping =
                       aes(x = Diet, y = Survived, shape = Diet, color = Diet)) + 
  stat_summary(colour = "red") +
  facet_wrap(~Temperature) +
   scale_fill_manual(values  = c("cornsilk4","lightgray"))+
  scale_y_continuous(limits =c(0,1))+
  labs(
    #title ="Diet and Temperatures \neffect on Survival",
     y = "Survival" , x = "Treatment" )

merged_surv_plot

# Development time 
merged_dev_plot <- ggplot(data = merged_performance_manduca_data,
                     mapping =
                       aes(x = Diet, y = Development_Time, fill = Temperature)) + 
  geom_boxplot() +
    scale_fill_manual(values  = c("cornsilk4","lightgray"))+
  geom_jitter(alpha = 0.5) +
  facet_wrap(~Temperature) +
 scale_y_continuous(limits =c(0,100)) +
  labs(
    #title = "Diet and Temperatures effect \non Development Time",
     y = "Development Time(days)" , x = "Treatment" )

merged_dev_plot

# Growth Rate
merged_growth_rate_plot <- ggplot(data = merged_performance_manduca_data,
                     mapping =
                       aes(x = Diet, y = Growth_Rate, fill = Diet)) + 
  geom_boxplot() +
  geom_jitter(alpha = 0.5) +
 facet_wrap(~Temperature) +
   scale_fill_manual(values  = c("cornsilk4","lightgray")) +
labs(y = "Mass(g) / Day " , x = "Treatment" ) +
  scale_y_continuous(limits =c(0,0.4)) +
theme(axis.title.y = element_text(size = 11)) 

merged_growth_rate_plot


```

New combined plot

merged

```{r}
fit_plot <- plot_grid(merged_pm_plot,merged_dev_plot,merged_growth_rate_plot, labels=c("A", "B","C"), ncol = 3, nrow = 1)

fit_plot
```

```{r}
surv_plot <- plot_grid(merged_surv_plot, ncol = 1, nrow = 1)

surv_plot
```

\

#### 2 \| Color (lightness) median

\
Anova

```{r}
Anova(merged_color_lightness_lm)
```

Sidak

```{r}
## merged
merged_color_lightness_em <- emmeans(merged_color_lightness_lm, specs = ~ Diet + Temperature)
merged_color_lightness_tk_Temp_by_Diet <- cld(merged_color_lightness_em, adjust = "sidak") %>%
  mutate(.group = letters[as.numeric(.group)]) %>%
  ggplot(mapping = aes(x =  Diet
                         , y = emmean,
                       ymin = lower.CL,
                       ymax = upper.CL,
                       colour = factor(.group))) +
  geom_pointrange() +
  geom_text(mapping = aes(label = .group),
            y = rep(1,4)) +
scale_color_manual(values  = c("black","cornflowerblue","honeydew3","snow4")) +
  facet_grid(~Temperature)

merged_color_lightness_tk_Temp_by_Diet
```

Original plot

```{r}
## merged
merged_color_lmed_plot <- ggplot(data = merged_color_manduca_data,
                     mapping =
                       aes(x = Diet, y = Luminance_Median, fill = Temperature)) + 
  geom_boxplot() +
  geom_jitter(alpha = 0.5) +
  scale_y_continuous(limits =c(50,130))+
  facet_wrap(~Temperature) +
   scale_fill_manual(values  = c("snow4","lightgrey"))+
  #scale_y_continuous(limits =c(0,0.10)) +
  labs(y = "Lightness median" , x = "Treatment" )

merged_color_lmed_plot
```

#### 2.5 \| Color (lightness) mean

Anova

```{r}
Anova(merged_color_lightness_mean_lm)
```

Sidak

```{r}
## merged
merged_color_lightness_mean_em <- emmeans(merged_color_lightness_mean_lm, specs = ~ Diet + Temperature)
merged_color_lightness_mean_tk_Temp_by_Diet <- cld(merged_color_lightness_mean_em, adjust = "sidak") %>%
  mutate(.group = letters[as.numeric(.group)]) %>%
  ggplot(mapping = aes(x =  Diet
                         , y = emmean,
                       ymin = lower.CL,
                       ymax = upper.CL,
                       colour = factor(.group))) +
  geom_pointrange() +
  geom_text(mapping = aes(label = .group),
            y = rep(1,4)) +
scale_color_manual(values  = c("black","cornflowerblue","honeydew3","snow4")) +
  facet_grid(~Temperature)

merged_color_lightness_mean_tk_Temp_by_Diet
```

Original plot

```{r}
## merged
merged_color_lmean_plot <- ggplot(data = merged_color_manduca_data,
                     mapping =
                       aes(x = Diet, y = Luminance_Mean, fill = Temperature)) + 
  geom_boxplot() +
  geom_jitter(alpha = 0.5) +
  scale_y_continuous(limits =c(50,130))+
  facet_wrap(~Temperature) +
   scale_fill_manual(values  = c("snow4","lightgrey"))+
  #scale_y_continuous(limits =c(0,0.10)) +
  labs(y = "Lightness mean" , x = "Treatment" )

merged_color_lmean_plot
```
